### 2023-08-22 sections/1-introduction.tex
edits to intro and sieve algorithm

\section{Introduction}
\label{sec:introduction}
Often dubbed "the Big Data explosion", rapidly advancing technologies have enabled researchers to collect complex data on an unprecedented scale. 
The Sloan Digital Sky Survey (SDSS), for example, has collected near-infrared spectra of approximately 430,000 stars in approximately 9,000 dimensions~\cite{alam2015eleventh}. 
Add sentence about MaNGA here: 10,000 galaxies with 2,700 deg$^2$, resolution $R=2,000$ for $\lambda \in [360, 1000) nm$.

Biology has also seen an influx of large, high-dimensional datasets. Silva 18S contains ribosomal DNA sequences of approximately 2.25 
million genomes with aligned length of 50,000 letters. (Do we need to look at NR99??)

In many cases, the increase in data outpaces the rate of improvement in computing capacity predicted by Moore's Law~\cite{moore1965cramming}, indicating 
that computer hardware will not "catch up" to our computational needs in the near future. As a result, analyzing and learning from these large datasets 
requires better algorithms.

The ability to perform similarity search on these datasets enables a variety of applications, including recommendation and classification systems. 
Additionally, with the rise of large language models such as GPT-3~\cite{brown2020language}, similarity search can become a useful tool for 
understanding and verifying the validity of model outputs. 
As the sizes and dimensionalities of datasets have grown, efficient and accurate similarity search has become excessively challenging; 
state-of-the-art algorithms exhibit a steep recall versus speed tradeoff~\cite{ishaq2019clustered}.
In particular, k-nearest neighbors (kNN) search is one of the most pervasive supervised classification methods in use~\cite{fix1952discriminatory, cover1967nearest}.

Naive implementations of kNN, whose asymptotic complexity is a function of the dataset's cardinality, 
prove prohibitively slow for large datasets. While algorithms for kNN on large datasets exist, most of them are approximate and fail to 
exploit the inherent structure and redundancy often present in big datasets. Approximate search is sufficient for many applications, but the 
need for efficient, exact search remains.

Recent approaches to tackling the exponential growth of data include locality-sensitive hashing~\cite{bingham2001random}, 
clever indexing techniques such as the FM Index~\cite{ferragina2005indexing}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}. 
Entropy-scaling search is a paradigm for similarity search that exploits the inherent structure and redundancy often present in big datasets.
Importantly, as suggested by their name, entropy-scaling search algorithms have asymptotic complexity that scales with geometric properties of the dataset,
rather than its cardinality. In 2019, we introduced CHESS (CLustered Hierarchical Entropy-Scaling Search), which extended entropy-scaling $\rho$-nearest 
neighbors search to a hierarchical clustering approach. In this paper, we introduce CLAM (Clustered Learning of Approximate Manifolds), and CAKES (CLAM-Accelerated K-nearest-neighbor 
Entropy-scaling Search). CLAM, like CHESS, is a hierarchical clustering algorithm (figure out and explain improvements upon CHESS present in CLAM if there are any algorithmic ones)
that can facilitate a variety of algorithms for big data. Using the tree constructed by CLAM, CAKES builds upon the capabilities of CHESS to perform 
k-nearest neighbors search in addition to $\rho$-nearest neighbors search.

At some point, need to address the fact that exact search is important for small k and approximate 
is maybe more ideal for large k. Try to find citation for this.

This paper focuses on:
\begin{itemize}
    \item Improvements and extensions of CHESS.
    \item Accelerated $\rho$-nearest and k-nearest neighbors search.
    \item Proof for how cluster-radii fall as we descend along the tree.
    \item A variety of distance functions including euclidean, manhattan, cosine, jaccard, hamming, levenshtein, needleman-wunch, and any user-defined function.
\end{itemize}

Comparisons for:
\begin{itemize}
    \item Search time vs FAISS and FALCONN and other in ann-benchmarks suite.
\end{itemize}

>>>>

\section{Introduction}
\label{sec:introduction}
Often dubbed "the Big Data explosion," rapidly advancing technologies have enabled researchers to collect complex data on an unprecedented scale. 
This is especially true in the field of astronomy. 
The Sloan Digital Sky Survey (SDSS), for example, in its APOGEE dataset, has collected near-infrared spectra of approximately 430,000 stars in approximately 9,000 dimensions~\cite{alam2015eleventh}. 
Their MaNGA dataset also contains spectral measurements across the face of each of about 10,000 galaxies across 2,700 deg$^2$.

Biology has also seen an influx of large, high-dimensional datasets. Silva 18S contains ribosomal DNA sequences of approximately 2.25 
million genomes with aligned length of 50,000 letters. (TODO: Do we need to talk about NR99; it seems like more of the same but 
with removed redundancy; not sure if that is helpful to us...)

In many fields, data collection rates outpace the rate of computing performance improvements predicted by Moore's Law~\cite{moore1965cramming}, indicating 
that computer architecture will not "catch up" to our computational needs in the near future. As a result, analyzing and learning from these large datasets 
requires better algorithms.

Many researchers would like the ability to perform similarity search on these datasets. 
Similatity search enables a variety of applications, including recommendation and classification systems. 
Additionally, with the rise of large language models such as GPT-3~\cite{brown2020language}, similarity search can become a useful tool for 
understanding and verifying the validity of model outputs. 
As the sizes and dimensionalities of datasets have grown, however, efficient and accurate similarity search has become extremely challenging; 
even state-of-the-art algorithms exhibit a steep recall versus speed tradeoff~\cite{ishaq2019clustered}.

In particular, $k$-nearest neighbors (kNN) search is one of the most pervasive supervised classification methods in use~\cite{fix1952discriminatory, cover1967nearest}.
Naive implementations of kNN, whose asymptotic complexity is a function of the dataset's cardinality, 
prove prohibitively slow for large datasets. While algorithms for fast $k$-NN on large datasets exist, most of them are approximate and do not 
exploit the inherent structure and redundancy often present in big datasets. Approximate search is sufficient for many applications, but the 
need for efficient, \emph{exact} search remains.

Recent approaches to tackling the exponential growth of data include locality-sensitive hashing~\cite{bingham2001random}, 
clever indexing techniques such as the FM Index~\cite{ferragina2005indexing}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}. 
Entropy-scaling search is a paradigm for similarity search that exploits the inherent structure and redundancy often present in big datasets.
Importantly, as suggested by their name, entropy-scaling search algorithms have asymptotic complexity that scales with geometric properties (such as the 
\emph{metric entropy} and \emph{local fractal dimension}, defined in Section 2) of a dataset,
rather than its cardinality. In 2019, we introduced CHESS (Clustered Hierarchical Entropy-Scaling Search), which extended entropy-scaling $\rho$-nearest 
neighbors search to a hierarchical clustering approach. In this paper, we introduce CLAM (Clustered Learning of Approximate Manifolds), and CAKES (CLAM-Accelerated K-nearest-neighbor 
Entropy-scaling Search). CLAM, like CHESS, is a hierarchical clustering algorithm 
that can facilitate a variety of algorithms for big data. Using the tree constructed by CLAM, CAKES builds upon the capabilities of CHESS to perform 
$k$-nearest neighbors search in addition to $\rho$-nearest neighbors search.

This paper focuses on improvements to CHESS, accelerated $\rho$- and $k$-nearest neighbors search, proof of a geometric guarantee for clusters, 
and the implementation of a variety of distance functions, including Euclidean, Manhattan, Cosine, Jaccard, Hamming, Levenshtein, Needleman-Wunch, 
and any user-defined function. We also provide a comparison of CAKES to FAISS~\cite{johnson2019billion}, FALCONN~\cite{andoni2015practical}, and
other algorithms in the ann-benchmarks suite~\cite{wang2017ann}.

### 2023-08-22 sections/2-methods.tex
edits to intro and sieve algorithm

We define a \emph{distance function} $f : (\textbf{X}, \textbf{X}) \mapsto \mathbb{R}^+ \cup \{0\}$ as a function which, 
given two points $x, y \in \textbf{X}$, deterministically returns a non-negative real number. We require that the distance function 
be symmetric ($f(x, y) = f(y, x) \forall x, y \in \textbf{X}$) and that the distance between two points $x$ and $y$ be zero if and only if $x = y$. 
When, in addition to these constraints, that the distance function obeys
the triangle inequality, it is a \emph{distance metric}, and in such cases we can guarantee that our search algorithms have perfect recall. 
Choice of distance function varies by type of data. For example, with electromagnetic spectra, we use both 
Euclidean (L2) and Cosine distances. With biological or string data on the other hand, Hamming and Levenshtein distances are more appropriate.

@@

Some of our algorithms for $\rho$- and $k$-nearest neighbors search rely on the \emph{local fractal dimension} of some region of the dataset, 
which we define as: 
$$\text{log}_2 \left( \frac{|B_D(q, r_1)|}{|B_D(q, r_2)|} \right)$$
where $B_D(q, r)$ is the set of points contained in a ball of radius $r$ 
centered on a point $q$ in the dataset $D$; in this example, we compute fractal dimension for some radius $r_1$ and a smaller radius $r_2 = \frac{r_1}{2}$.
We stress that this concept of local fractal dimension differs from the \emph{embedding dimension} of a dataset. To illustrate the difference,
consider SDSS's APOGEE dataset, wherein each datum is a nonnegative real-valued vector of length 8,575. Hence, the \emph{embedding dimension} of this dataset is 8,575. 
However, due to physical constraints (namely, the laws that govern stellar fusion and spectral emission lines), the data are constrained to a lower-dimensional 
manifold. The \emph{local fractal dimension} is an approximation of the dimensionality of that lower-dimensional manifold at a given point, for some length scale.
The notion that high-dimensional data collected from constrained generating phenomena typically only occupy a low-dimensional manifold is known as the \emph{manifold hypothesis}.

@@

We sometimes also use the above notation with points instead of clusters. For example, we sometimes use $\delta = f(x, q)$ to refer to the distance 
from a point $x$ to the query $q$. Since we can think of a point as having radius 0, for points we have that $\delta_{max} = \delta = \delta_{min}$.

@@

We start by building a divisive hierarchical clustering of the dataset with CLAM, using a 
similar recursive procedure as outlined in CHESS~\cite{ishaq2019clustered}, but with the following 
improvements: better selection of poles for partitioning and memoization of useful cluster properties, 
such as local fractal dimension (see Section ADD SECTION: $k$-NN by Repeated $\rho$-NN).

@@

For a cluster with $m$ points, we begin by computing a 
random sample of $\sqrt m$ points. We then compute the geometric median of this sample, which we call the 
cluster's center. We then compute the distance from the center to each point in the sample of $\sqrt m$ points. 
The point $l$ which is furthest from the center is designated the left pole, and the point $r$ which is furthest
from $l$ is designated the right pole. We then partition the cluster into a left child and a right child, where the 
left child contains all points in the cluster which are closer to $l$ than to $r$, and the right child contains all 
points in the cluster which are closer to $r$ than to $l$. Without loss of generality, we assign to the left child 
those points which are equidistant from $l$ and $r$. Starting from a root-cluster containing the entire dataset, we 
repeat this procedure until each leaf contains only one datum or until some other user-specified stopping criterion 
is met.

@@

Assume that we have a dataset which follows a $d$-dimensional distribution embedded in $\mathcal{D}$-dimensional space.
We can describe this distribution choosing some set of $d$ mutually orthogonal axes.
Let $m$ denote the maximum distance between any pair of points along any axis from any choice of mutually orthogonal axes for our $d$-dimensional distribution.
In the worst case scenario, the distribution of our data is a $d$-sphere and $m$ is the maximum distance between two points along every axis for every choice of axes.
In this scenario, the cluster radius is $r = \frac{m}{2}$.

Given a cluster $\mathcal{C}$, our partition algorithm will choose the left and right poles, 
and the axis defined by those poles will be the axis for partitioning $\mathcal{C}$'s points into two child clusters.
For points in each child cluster, the maximum distance between any pair of points along the \emph{axis of partitioning} will have been reduced to at most $\frac{m}{2}$.
Though this bound holds for points on the axis of partitioning, it does not serve as a bound for the maximum distance between \emph{any} pair of points in $\mathcal{C}$.
To see this, consider the two dimensional scenario where the points in $\mathcal{C}$ are distributed in a circle of radius $r$. Clearly, the maximum possible distance between 
any two points in $\mathcal{C}$ is $2r$, and occurs when the points fall on opposite ends of one of its diameters. Now consider a horizontal diameter of the circle, 
and let $l$ denote the leftmost point on this diameter. Next, consider the orthogonal vertical diameter, and let $x$ denote the uppermost point on this diameter. When we 
partition along the horizontal diameter, $l$ and $x$ will be assigned to the left child (as mentioned in ADD SECTION, we assign equidistant points to the left child). 
This reduces the maximum possible distance between any two points on the axis of partition -- the horizontal diameter -- to $r$. However, by considering the 
right triangle formed by $l$, $x$, and the center of the circle, we can see that the distance between 
$l$ and $x$ is $\sqrt{2}{r}$.

@@

Given a query $q$ and a search radius $\rho$, find all $x \in X$ s.t. $f(q, x) \leq \rho$.

We conduct ranged-nearest neighbors search similarly as described in CHESS~\cite{ishaq2019clustered}, but 
with some implementation improvements including identification of candidate clusters which are leaves and 
conducting exhaustive search over those leaves.

@@

In this section, we present four novel variants of $k$-nearest neighbors search: Greedy Search, K-NN by Repeated $\rho$-NN, Sieve Search I, and Sieve Search II. 
We use a preliminary auto-tuning function (TODO: do we though?) to predict the variant which will perform 
best for a given query, dataset, and value of $k$. We then proceed with search using that variant.

\subsubsection{Greedy Search}
\label{subsubsec:methods:knn-search:greedy-search}
Greedy Search repeatedly partitions the cluster with the lowest $\delta_{min}$ into its children 
until it is a leaf, removing it and adding its points to a fixed size priority queue of points. Search terminates 
when the furthest hit in the queue of points has $\delta$ greater than the $\delta_{min}$ of the closest remaining
cluster.

@@

While $P$ has fewer than $k$ clusters or $\delta$ of the top priority element in $P$ is greater 
than $\delta_{min}$ of the top priority element in $Q$, do the following:
\begin{enumerate}
\item While the top priority element $q$ in $Q$ isn't a leaf, replace $q$ with its children.
\item Remove $q$ from $Q$ and add all of its points to $P$. 
\item Remove points from $P$ until $P$ has $k$ points. 
\end{enumerate}

Return $P$.

\begin{algorithm} 
\caption{GreedySearch(\emph{tree, query, k})} 
\label{alg:greedy_search} 
\begin{algorithmic}[3]
    \STATE $Q \leftarrow$ priority queue
    \STATE $Q.push(tree.root)$
    \STATE $P \leftarrow$ priority queue
    \WHILE{$|P| < k \lor P.max.\delta > Q.min.\delta_{min}$}
        \WHILE{$!Q.min.isLeaf$}
            \STATE $[l, r] \leftarrow Q.extractMin().children()$
            \STATE $Q.push([l, r])$
        \ENDWHILE
        \STATE $leaf \leftarrow Q.extractMin()$
        \STATE $P.push(leaf)$
        \WHILE{$|P| > k$}
            \STATE $P.extractMax()$
        \ENDWHILE
    \ENDWHILE
    \STATE Return $P$
\end{algorithmic}
\end{algorithm}

\subsubsection{$k$-NN by Repeated $\rho$-NN}
\label{subsubsec:methods:knn-search:repeated-rnn}
Hence the name, with $k$-NN by Repeated $\rho$-NN we repeatedly perform
$\rho$-nearest neighbors search, increasing the search radius until $k$ neighbors
have been found.

@@

We start by letting $Q$ be a vector of clusters and points sorted by non-decreasing $\delta_{max}$ and initializing $Q$ with the root cluster. 
While we haven't found a threshold which guarantees exactly $k$ points, we do the following:

TODO: Fix this to reflect new algorithm.

\begin{enumerate}
\item Find the cluster $C_{\tau}$ in $Q$ with the smallest $\delta_{max}$ such that no cluster whose $\delta_{min}$ is greater than $C_{\tau}$'s $\delta_{max}$ can contain one of the $k$ nearest neighbors.
\item Let the threshold $\tau$ be the $\delta_{max}$ of $C_{\tau}$.
\item Remove from $Q$ any cluster whose $\delta_{min}$ is greater than $\tau$.
\item Transfer from $Q$ to $P$ all the points in any cluster whose cardinality is less than $k$ or who is a leaf. 
\item Replace all non-leaf clusters in $Q$ with their children. 
\end{enumerate}

At this stage, if any clusters remain in $Q$, we add all their points to $P$. 
Extract $P$.

\begin{algorithm} % enter the algorithm environment
    \caption{Sieve } % give the algorithm a caption
    \label{alg:sieve_v1} % and a label for \ref{} commands later in the document
    \begin{algorithmic}[2] % enter the algorithmic environment
        \REQUIRE $tree$, $query$, $k$
        \STATE $Q \leftarrow$ priority queue
        \STATE $Q.push(tree.root)$
        \STATE $P \leftarrow$ priority queue
        \WHILE{$|P| < k \lor P.max.\delta > Q.max.\delta_{min}$}
            \WHILE{$!Q.extractMax().isLeaf$}
                \STATE $[l, r] \leftarrow Q.extractMax().children()$
                \STATE $Q.push([l, r])$
            \ENDWHILE
            \STATE $leaf \leftarrow Q.extractMax()$
            \STATE $P.push(leaf)$
            \WHILE{$|P| > k$}
                \STATE $P.extractMax()$
            \ENDWHILE
        \ENDWHILE
        \STATE Return $P$
    \end{algorithmic}
    \end{algorithm}

@@

\subsubsection{Leaf Search}
\label{subsubsec:methods:knn-search:leaf-search}

>>>>

We define a \emph{distance function} $f : (\textbf{X}, \textbf{X}) \mapsto \mathbb{R}^+ \cup \{0\}$ as a function which, 
given two points $x, y \in \textbf{X}$, deterministically returns a non-negative real number. We require that the distance function 
be symmetric ($f(x, y) = f(y, x)$ for all $x, y \in \textbf{X}$) and that the distance between two points $x$ and $y$ be zero if and only if $x = y$. 
When, in addition to these constraints, the distance function obeys
the triangle inequality, it is a \emph{distance metric}, and in such cases we can guarantee that our search algorithms have perfect recall. 
Choice of distance function varies by type of data. For example, with electromagnetic spectra, we use both 
Euclidean (L2) and Cosine distances. With biological or string data on the other hand, Hamming and Levenshtein distances are more appropriate.

Some of our algorithms for $\rho$- and $k$-nearest neighbors search rely on the \emph{local fractal dimension} at some point in the dataset, 
which we define as: 
$$\text{log}_2 \left( \frac{|B_X(q, r_1)|}{|B_X(q, r_2)|} \right)$$
where $B_X(q, r)$ is the set of points contained in a ball of radius $r$ 
centered at a point $q$ in the dataset $\textbf{X}$; in this example, we compute fractal dimension for some radius $r_1$ and a smaller radius $r_2 = \frac{r_1}{2}$.
We stress that this concept differs from the \emph{embedding dimension} of a dataset. To illustrate the difference,
consider SDSS's APOGEE dataset, wherein each datum is a nonnegative real-valued vector of length 8,575. Hence, the \emph{embedding dimension} of this dataset is 8,575. 
However, due to physical constraints (namely, the laws that govern stellar fusion and spectral emission lines), the data are constrained to a lower-dimensional 
manifold within the 8,575-dimensional embedding space. The \emph{local fractal dimension} is an approximation of the dimensionality of that lower-dimensional manifold at a given point, for some length scale.
The notion that high-dimensional data collected from constrained generating phenomena typically only occupy a low-dimensional manifold is known as the \emph{manifold hypothesis}.

@@

@@

We sometimes also use this notation with points instead of clusters. For example, we sometimes use $\delta = f(x, q)$ to refer to the distance 
from a point $x$ to the query $q$. Since we can think of a point as having radius $r=0$, for points we have that $\delta_{max} = \delta = \delta_{min}$.

@@

We start by building a divisive hierarchical clustering of the dataset with CLAM, using a 
similar recursive procedure as outlined in CHESS~\cite{ishaq2019clustered}, but with the following 
improvements: better selection of poles for partitioning and memoization of useful cluster properties, 
such as local fractal dimension (see Section 2.4.2: $k$-NN by Repeated $\rho$-NN).

@@

For a cluster with $m$ points, we begin by computing a 
random sample of $\sqrt m$ points. We then compute the geometric median of this sample, which we call the 
cluster's center. We then compute the distance from the center to each point in the sample of $\sqrt m$ points. 
The point $l$ which is furthest from the center is designated the left pole, and the point $r$ which is furthest
from $l$ is designated the right pole. We then partition the cluster into a left child and right child, where the 
left child contains all points in the cluster which are closer to $l$ than to $r$, and the right child contains all 
points in the cluster which are closer to $r$ than to $l$. Without loss of generality, we assign to the left child 
those points which are equidistant from $l$ and $r$. Starting from a root-cluster containing the entire dataset, we 
repeat this procedure until each leaf contains only one datum or until some other user-specified stopping criterion 
is met.

@@

Assume that we have a dataset which follows a $d$-dimensional distribution embedded in $D$-dimensional space.
We can describe this distribution choosing some set of $d$ mutually orthogonal axes.
Let $m$ denote the maximum distance between any pair of points along any axis from any choice of mutually orthogonal axes for our $d$-dimensional distribution.
In the worst case scenario, the distribution of our data is a $d$-sphere and $m$ is the maximum distance between two points along every axis for every choice of axes.
In this scenario, the cluster radius is $r = \frac{m}{2}$.

Given a cluster $C$, we choose the left and right poles, 
and the axis defined by those poles will be the axis for partitioning  $C$'s points into two child clusters.
For points in each child cluster, the maximum distance between any pair of points along the \emph{axis of partitioning} will have been reduced to at most $\frac{m}{2}$.
Though this bound holds for points on the axis of partitioning, it does not serve as a bound for the maximum distance between \emph{any} pair of points in  $C$.
To see this, consider the two dimensional scenario where the points in  $C$ are distributed in a circle of radius $r$. Clearly, the maximum possible distance between 
any two points in  $C$ is $2r$, and occurs when the points fall on opposite ends of one of its diameters. Now consider a horizontal diameter of the circle, 
and let $l$ denote the leftmost point on this diameter. Next, consider the orthogonal vertical diameter, and let $x$ denote the uppermost point on this diameter. When we 
partition along the horizontal diameter, $l$ and $x$ will be assigned to the left child (as mentioned in Section 2, we assign equidistant points to the left child). 
This reduces the maximum possible distance between any two points on the axis of partition --the horizontal diameter-- to $r$. However, by considering the 
right triangle formed by $l$, $x$, and the center of the circle, we can see that the distance between 
$l$ and $x$ is $\sqrt{2}{r}$.

@@

We conduct $\rho$-nearest neighbors search similarly as described in CHESS~\cite{ishaq2019clustered}, but 
with some implementation improvements including identification of candidate clusters which are leaves and 
conducting exhaustive search over those leaves.

@@

In this section, we present four novel variants of $k$-nearest neighbors search: Greedy Search, $k$-NN by Repeated $\rho$-NN, Sieve Search I, and Sieve Search II. 
We use a preliminary auto-tuning function (TODO: do we though?) to predict the variant which will perform 
best for a given query, dataset, and value of $k$. We then proceed with search using that variant.

\subsubsection{Greedy Search}
\label{subsubsec:methods:knn-search:greedy-search}
Greedy Search repeatedly partitions the cluster with the smallest $\delta_{min}$ into its children 
until the cluster with the smallest $\delta_{min}$ is a leaf. We then remove it and add its points to a fixed size priority queue of points. 
Search terminates when the furthest hit in the queue of points has $\delta$ greater than the $\delta_{min}$ of the closest remaining cluster.

@@

While $P$ has fewer than $k$ clusters or $\delta$ of the top priority element in $P$ is greater 
than $\delta_{min}$ of the top priority element in $Q$, do the following:
\begin{enumerate}
\item While the top priority element $q$ in $Q$ is not a leaf, replace $q$ with its children.
\item Remove $q$ from $Q$ and add all of its points to $P$. 
\item Remove points from $P$ until $P$ has $k$ points. 
\end{enumerate}
After this process, the points left in $P$ are the $k$ nearest neighbors of $q$.

\begin{algorithm} 
\caption{GreedySearch(\emph{tree, query, k})} 
\label{alg:greedy_search} 
\begin{algorithmic}[3]
    \STATE $Q \leftarrow$ priority queue
    \STATE $Q.push(tree.root)$
    \STATE $P \leftarrow$ priority queue
    \WHILE{$|P| < k \lor P.max.\delta > Q.min.\delta_{min}$}
        \WHILE{$\neg Q.min.isLeaf$}
            \STATE $[l, r] \leftarrow Q.extractMin().children$
            \STATE $Q.push([l, r])$
        \ENDWHILE
        \STATE $leaf \leftarrow Q.extractMin()$
        \STATE $P.push(leaf)$
        \WHILE{$|P| > k$}
            \STATE $P.extractMax()$
        \ENDWHILE
    \ENDWHILE
    \STATE Return $P$
\end{algorithmic}
\end{algorithm}

\subsubsection{$k$-NN by Repeated $\rho$-NN}
\label{subsubsec:methods:knn-search:repeated-rnn}
In this $k$-NN variant, we repeatedly perform
$\rho$-nearest neighbors search, increasing the search radius until $k$ neighbors
have been found.

@@

We start by letting $Q$ be a vector of clusters and points sorted by non-decreasing $\delta_{max}$ and initializing $Q$ with the root cluster. 
While we haven't found a threshold which guarantees exactly $k$ points, we repeat the following:

\begin{enumerate}
\item Find the cluster $C_{\tau}$ in $Q$ with the smallest $\delta_{max}$ such that no cluster whose $\delta_{min}$ is greater than $C_{\tau}$'s $\delta_{max}$ can contain one of the $k$ nearest neighbors.
\item Let the threshold $\tau$ be the $\delta_{max}$ of $C_{\tau}$.
\item Remove from $Q$ any cluster whose $\delta_{min}$ is greater than $\tau$.
\item For each cluster left in $Q$, if it is a leaf or has $k$ or fewer points, remove it and all of its points to $Q$.
\item Replace all remaining non-leaf clusters in $Q$ with their children. 
\end{enumerate}

The points remaining in $Q$ are the $k$ nearest neighbors.

TODO: Add find kth algorithm.

\begin{algorithm} % enter the algorithm environment
    \caption{Sieve(\emph{tree, query, k})} % give the algorithm a caption
    \label{alg:sieve1} % and a label for \ref{} commands later in the document
    \begin{algorithmic}[5] % enter the algorithmic environment
        \REQUIRE $tree$, $query$, $k$
        \STATE $Q \leftarrow$ priority queue
        \STATE $Q.push(tree.root)$
        \WHILE{$|Q| > k$}
            \STATE $C_{\tau} \leftarrow Q.findKth.\delta_{max}$
            \FOR {$q \in Q$}
                \IF {$q.\delta_{min} > \tau$}
                    \STATE $Q.remove(q)$
                \ENDIF
            \ENDFOR
            \FOR {$q \in Q$}
                \IF {$q.isLeaf \lor |q.cardinality| \leq k$}
                    \STATE $Q.push(q.points)$
                \ELSE
                    \STATE $[l, r] \leftarrow q.children$
                    \STATE $Q.push([l, r])$   
                \ENDIF
                \STATE $Q.remove(q)$
            \ENDFOR 
        \ENDWHILE
        \STATE Return $Q$
    \end{algorithmic}
    \end{algorithm}

@@
